---
title: "CourseProjectRMD"
author: "Fahsi"
date: "11 12 2023"
output: html_document
---



# Executive summary
# Analysis
## Loading data
Loading the data and splitting it into training and testing Datasets
```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
csv_train <- '/home/rstudio/Project/Proj/pml-training.csv'
csv_test <- '/home/rstudio/Project/Proj/pml-testing.csv'

training <- read.csv(file=csv_train,sep=',',header=T,na.strings=c("NA","NaN", "#DIV/0!"))
testing<- read.csv(file=csv_test,sep=',',header=T,na.strings=c("NA","NaN", "#DIV/0!"))
library(caret) 
library(randomForest)
set.seed(13234)
testIndex = createDataPartition(training$classe, p = 0.30,list=FALSE)
train = training[-testIndex,]
test = training[testIndex,]
```

## Dealing with missing values (sparse variables) 
Detecting and ignoring columns that do not fit logicaly as Predictors 
```{r}
na_values <- is.na(train)
na_means<-colMeans(na_values)
c <- (colMeans(na_values) < min(na_means[na_means>0]))
train_1 <- train[,c]
summary(c) 
#New Training Data Set with 60 Columns  
```

  
## Omitting sample specific variables
droping Datafield that are not appropriate as predictors (time stamps, user name, ...)
```{r}
drop <-  c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp" ,"new_window","num_window")
train_2 <- train_1[,!(names(train_1) %in% drop)]
dim(train_2)
```
#Reduced dimensions.

## Choosing a classification Model for prediction for this type of data 
Training Data with Random Forest Model 
```{r}
train_2$classe <- as.factor(train_2$classe)
 
response_variable <- "classe"
predictors <- setdiff(names(train_2), response_variable)

#Creating Random Forest Modell
rf_model_1<-randomForest(classe ~ .,data=train_2)

#Summary of the created model
print(rf_model_1)
```
#Calculating Variable Significance
#library(caret)

```{r}
imp <- varImp(rf_model_1)
summary(imp)
```

#New Data set without Variables with lower significance 
```{r}
low_signific<-imp<95
names(train_2[,low_signific])
```

#Droping Data with low Significance

```{r}
#Geo Coordinates data
kuerzel_liste <- c("_x", "_y", "_z")

# Drop columns containing _x, _y , _z
train_3 <- train_2[, !grepl(paste(kuerzel_liste, collapse = "|"), names(train_2))]

dim(train_3)
#Other data with low significance 
drop <-  c("total_accel_forearm", "total_accel_arm" )
train_3 <- train_3[,!(names(train_3) %in% drop)]
dim(train_3)
```

#Trying Random Forest Model with the new reduced training data 
```{r}
library(randomForest)
library(caret)

set.seed(123)

# Define the training control for cross-validation
ctrl <- trainControl(
  method = "cv",        # Cross-validation
  number = 5,           # Number of folds
  verboseIter = TRUE    # Print progress
)

# Specify the formula for the model
formula <- classe ~ .

# Train the Random Forest model with cross-validation
rf_model_2 <- train(
  formula,
  data = train_3,
  method = "rf",
  trControl = ctrl
)

# Print the results
print(rf_model_2)

# Access variable importance
var_importance <- varImp(rf_model_2)
print("Variable Importance:")
print(var_importance)

train_3$classe <- as.factor(train_3$classe)
test$classe<- as.factor(test$classe)
# Make predictions (replace "YourNewData" with your new data)
# For example purposes, we'll use the same iris dataset
predictions <- predict(rf_model_2, newdata = test)

# Evaluate the model
conf_matrix <- confusionMatrix(predictions, test$classe)
print("Confusion Matrix:")
print(conf_matrix)


# Summary of additional metrics
summary(conf_matrix)
```

# Plotting the confusion matrix with ggplot
```{r}
# Plotting the confusion matrix with ggplot
conf_matrix_data <- as.data.frame(as.table(conf_matrix))

# Plot the confusion matrix with ggplot2
ggplot(conf_matrix_data, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%d", Freq)), vjust = 1) +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_minimal() +
  labs(x = "Actual", y = "Predicted", fill = "Frequency")
```

## Predicting on the testing set using the Random Forest Model

```{r}
predict(rf_model_2,testing)
```


